---
title: US flight cancellations in 2008
author: Gautam Joshi
date: '2017-12-12'
slug: us-flight-cancellations-in-2008
categories:
  - Exploratory Data Analysis
  - d3.js
tags:
  - EDA
  - d3.js
---


<body style = "font-family:Georgia">

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F)
```


```{r}
# libraries needed
library(tidyverse)
library(formattable)
library(RColorBrewer)
library(stringr)
library(DT)
library(scales)
library(gridExtra)
```


```{r}
# set directory
setwd("/Users/gautamjoshi/Desktop/R_std")
# load data
flights <- read_csv('2008.csv')
```


Sample Data


```{r}
flights %>%
  head(10) %>%
  formattable(align = "l") %>% 
  datatable(options = list(scrollX = "400px"))
```



Updating column Month with month names and DayOfWeek with day names.

```{r echo=T}
newflights <- flights

newflights$DayOfWeek[newflights$DayOfWeek == 1] = 'Monday'
newflights$DayOfWeek[newflights$DayOfWeek == 2] = 'Tuesday'
newflights$DayOfWeek[newflights$DayOfWeek == 3] = 'Wednesday'
newflights$DayOfWeek[newflights$DayOfWeek == 4] = 'Thursday'
newflights$DayOfWeek[newflights$DayOfWeek == 5] = 'Friday'
newflights$DayOfWeek[newflights$DayOfWeek == 6] = 'Saturday'
newflights$DayOfWeek[newflights$DayOfWeek == 7] = 'Sunday'

newflights$Month[newflights$Month == 1] = 'January'
newflights$Month[newflights$Month == 2] = 'February'
newflights$Month[newflights$Month == 3] = 'March'
newflights$Month[newflights$Month == 4] = 'April'
newflights$Month[newflights$Month == 5] = 'May'
newflights$Month[newflights$Month == 6] = 'June'
newflights$Month[newflights$Month == 7] = 'July'
newflights$Month[newflights$Month == 8] = 'August'
newflights$Month[newflights$Month == 9] = 'September'
newflights$Month[newflights$Month == 10] =  'October'
newflights$Month[newflights$Month == 11] =  'November'
newflights$Month[newflights$Month == 12] =  'December'
```

Creating a dataframe containing count of cancelled flights grouped by Months and Days of Week. Also, arranging the dataframe in chronological order.

```{r echo=T}
cancelledByMonthDay <- newflights %>%
  filter(Cancelled == 1) %>%
  group_by(Month, DayOfWeek) %>%
  count()

cancelledByMonthDay$Month <- factor(cancelledByMonthDay$Month, 
                                        levels = month.name)
cancelledByMonthDay$DayOfWeek <- factor(cancelledByMonthDay$DayOfWeek, 
                                           levels = c("Sunday", 
                                                      "Monday", 
                                                      "Tuesday", 
                                                      "Wednesday", 
                                                      "Thursday", 
                                                      "Friday", 
                                                      "Saturday"))

colnames(cancelledByMonthDay)[3] <- "Count"

cancelledByMonthDay <- cancelledByMonthDay[order(cancelledByMonthDay$DayOfWeek),]
cancelledByMonthDay <- arrange(cancelledByMonthDay, Month)

cancelledByMonthDay %>%
  datatable()
```

Plotting grouped bar chart to display total annual cancellations separated by months and days of week. Furthermore, this plot is an intital sketch of the final plot that I intend to create with D3.js. 

```{r}
cancelledByMonthDay %>%
  ggplot(aes(x = Month, y = Count)) +
  geom_bar(aes(fill = DayOfWeek), stat = "summary",
           position = "dodge",
           color = "white",
           width = 0.7,
           size = 0.3) +
  ylab("Cancellations") +
  ggtitle("Total Cancelled flights in 2008 by Months and Days of Week") +
  theme_minimal() +
  guides(fill=guide_legend(title="Day Of Week")) +
  theme(axis.text.x = element_text(size = 6, 
                                   angle = 45),
        axis.text.y = element_text(size = 6),
        plot.title = element_text(family = "Georgia", 
                                  size = 10,
                                  color = "darkgrey",
                                  face = "bold",
                                  hjust = 0.5),
        legend.key.size = unit(.2, "cm"),
        legend.title = element_text(size = 6),
        legend.text = element_text(size = 4),
        legend.position = c(0.92, 0.9)) +
  scale_fill_brewer(palette = "YlGnBu")
```


Reshaping dataframe for D3.js ingestion.  


```{r echo=T}
cancelledByMonthDay$Sunday <- cancelledByMonthDay$Count[cancelledByMonthDay$DayOfWeek == "Sunday"]
cancelledByMonthDay$Monday <- cancelledByMonthDay$Count[cancelledByMonthDay$DayOfWeek == "Monday"]
cancelledByMonthDay$Tuesday <- cancelledByMonthDay$Count[cancelledByMonthDay$DayOfWeek == "Tuesday"]
cancelledByMonthDay$Wednesday <- cancelledByMonthDay$Count[cancelledByMonthDay$DayOfWeek == "Wednesday"]
cancelledByMonthDay$Thursday <- cancelledByMonthDay$Count[cancelledByMonthDay$DayOfWeek == "Thursday"]
cancelledByMonthDay$Friday <- cancelledByMonthDay$Count[cancelledByMonthDay$DayOfWeek == "Friday"]
cancelledByMonthDay$Saturday <- cancelledByMonthDay$Count[cancelledByMonthDay$DayOfWeek == "Saturday"]
cancelledByMonthDay$Month <- unique(cancelledByMonthDay$Month)
cancelledByMonthDay$DayOfWeek <- NULL
cancelledByMonthDay$Count <- NULL
cancelled_df <- unique(cancelledByMonthDay)
```


Writing the reshaped dataframe to csv.

```{r echo=T}
write_csv(cancelled_df, "total_cancellations_2008.csv")
```

</body>
